---
title: "Product Demand Forecast-Monthly By Node Group Model (hts)"
author: "Microsoft"
output: 
    rmarkdown::html_vignette:
        toc: true

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```

# Introduction
In this documentation, we are trying to build a monthly by node group demand forecast model on product booking quantity data using the package 'hts'. We implement the advanced analysis process with R in the principle of step-by-step. 

## Product booking quantity data
Let's load the data and take a brief glimpse of it. 
```{r, message=FALSE}
##########################################################################
## Set directory
##########################################################################
setwd("C:/Users/zhouf/Documents/Rproject")
data.path <- "C:/Users/zhouf/Documents/Rproject"
```

```{r, message=FALSE, warning=FALSE, error=FALSE}
##########################################################################
## Read data
##########################################################################
data <- read.table("Data/After scramble-Node group.csv", header=TRUE, sep=",")
dim(data) # 120 17
str(data)
head(data)
```

## Handling missing value
To clean the data, we remove some of columns with too many NAs (>10%) and replace other missing values with last-observation-carried-forward method. 
```{r, message=FALSE, warning=FALSE, error=FALSE}
##########################################################################
## Clean missing value
##########################################################################
library(plyr)

# remove columns with only NA
data <- Filter(function(x)!all(is.na(x)), data)
dim(data) # 120 16

# remove columns with too many NAs 
data <- Filter(function(x)!(sum(is.na(x))/length(x)) > 0.1, data)
dim(data) # 120 10

# remove rows with only missing values
missingRows <- apply(data, 1, function(x) all(is.na(x)))
data <- data[!missingRows, ]
dim(data)

# replace NA values with last-observation-carried-forward method
library(zoo)
data[,3:10] <- lapply(data[,3:10], function(x) { 
  x <- na.locf(x)
  x
})
head(data)
```

## Model building
We aim to build a time series model to fit the whole product booking quantity data. 

### Hierachical time series model
Hierarchical time series forecasting methods can be implemented in the hts package (Hyndman et al. 2014) for R (R Core Team 2013).

We can use hts function to create a hierarchical time series. The required inputs are the bottom-level time series, and information about the hierarchical structure. Here, we try two hierarchical structure: first, each bottom series belong to a group; second, the series for product 1,3,4,5,6 belong to one group, series for product 8,9,10 belong to the second group, and other bottom series belong to the third group.

Let's have a look at 2-level hierarchical time series. 

```{r, message=FALSE, warning=FALSE, error=FALSE}
##########################################################################
## Hierarchical time series modeling 
##########################################################################
library(hts)

# construct time series matrix
bts <- ts(data[, 4:10], frequency=12, start=c(2006, 1))
head(bts)

# create 2-level hierarchical time series
demandgts <- hts(bts)
print(demandgts)
alldemand <- aggts(demandgts) 
head(alldemand)
plot(demandgts, levels=c(0,1))

# forecast 3-step-ahead demand using the bottom-up method
demandforecast <- forecast(demandgts, h=3, method="bu", fmethod="arima", 
                           keep.fitted = TRUE)
# plot the forecasts inlcluding the last ten historical years (120 historical months)
plot(demandforecast, include=120)
aggts(demandforecast)
summary(demandforecast)

# set up the training sample and testing sample
train <- window(demandgts, start=c(2006, 1), end=c(2014, 12))
test <- window(demandgts, start=c(2015, 1), end=c(2015, 12))
demandforecast <- forecast(train, h=12, method="bu", fmethod="arima")
# calculate perforemance metrics
accuracy.gts(demandforecast, test)
aggts(demandforecast)
plot(demandforecast)
```

Let's try a 3-level hierarchical time series.

```{r, message=FALSE, warning=FALSE, error=FALSE}
# put the series within the same group together and construct new time series matrix
newdata <- data[, -5]
newdata$Product.2 <- data[, 5]
head(newdata)
bts <- ts(newdata[, 4:10], frequency=12, start=c(2006, 1))
head(bts)

# create multi-level hierarchical time seris
demandgts <- hts(bts, nodes=list(3, c(4,2,1)))
print(demandgts)
alldemand <- aggts(demandgts) 
head(alldemand)
plot(demandgts)

# forecast 3-step-ahead demand using the bottom-up method
demandforecast <- forecast(demandgts, h=3, method="bu", fmethod="arima", 
                           keep.fitted=TRUE)
# plot the forecasts inlcluding the last ten historical years (120 historical months)
plot(demandforecast, include=120)
aggts(demandforecast)
summary(demandforecast)

# set up the training sample and testing sample
train <- window(demandgts, start=c(2006, 1), end=c(2014, 12))
test <- window(demandgts, start=c(2015, 1), end=c(2015, 12))
demandforecast <- forecast(train, h=12, method="bu", fmethod="arima", keep.fitted=TRUE)
summary(demandforecast)
# calculate perforemance metrics
accuracy.gts(demandforecast, test)
aggts(demandforecast)
plot(demandforecast)
``` 

From a model performance metrics perspective, there seems to be no much difference between 2-level and 3-level hierarchical time series model.

### Cross-validation for hierarchical time series

Next, we use cross validation to compare performance of different forecast methods for hierarchical time series model.

```{r, message=FALSE, warning=FALSE, error=FALSE}
# use time series cross validation to compare model performance
library(fpp)

k <- 96 # minimum data length for fitting a model
n <- dim(demandgts$bts)[1]
st <- tsp(demandgts$bts)[1]+(k-2)/12
mape11 <- mape12 <- mape13 <- mape21 <- mape22 <- mape23 <- mape31 <- mape32 <- mape33 <- matrix(NA,n-k,12)
for(i in 1:(n-k))
{
  xshort <- window(demandgts, end=st + i/12)
  xnext <- window(demandgts, start=st + (i+1)/12, end=st + (i+12)/12)
  fcast11 <- forecast(xshort, h=12, method="bu", fmethod="arima", keep.fitted = TRUE)
  fcast12 <- forecast(xshort, h=12, method="bu", fmethod="ets", keep.fitted=TRUE)
  fcast13 <- forecast(xshort, h=12, method="bu", fmethod="rw", keep.fitted=TRUE)
  fcast21 <- forecast(xshort, h=12, method="comb", fmethod="arima", keep.fitted = TRUE)
  fcast22 <- forecast(xshort, h=12, method="comb", fmethod="ets", keep.fitted=TRUE)
  fcast23 <- forecast(xshort, h=12, method="comb", fmethod="rw", keep.fitted=TRUE)
  fcast31 <- forecast(xshort, h=12, method="tdgsa", fmethod="arima", keep.fitted = TRUE)
  fcast32 <- forecast(xshort, h=12, method="tdgsa", fmethod="ets", keep.fitted=TRUE)
  fcast33 <- forecast(xshort, h=12, method="tdgsa", fmethod="rw", keep.fitted=TRUE)
  mape11[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast11), aggts(xnext))[1, "MAPE"]
  mape12[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast12), aggts(xnext))[1, "MAPE"]
  mape13[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast13), aggts(xnext))[1, "MAPE"]
  mape21[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast21), aggts(xnext))[1, "MAPE"]
  mape22[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast22), aggts(xnext))[1, "MAPE"]
  mape23[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast23), aggts(xnext))[1, "MAPE"]
  mape31[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast31), aggts(xnext))[1, "MAPE"]
  mape32[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast32), aggts(xnext))[1, "MAPE"]
  mape33[i,1:dim(xnext$bts)[1]] <- accuracy(aggts(fcast33), aggts(xnext))[1, "MAPE"]
}

plot(1:12, colMeans(mape11,na.rm=TRUE), type="l", col=2, 
     xlab="horizon", ylab="Overall MAPE", ylim=c(10, 25))
lines(1:12, colMeans(mape12,na.rm=TRUE), type="l",col=3)
lines(1:12, colMeans(mape13,na.rm=TRUE), type="l",col=4)
lines(1:12, colMeans(mape21,na.rm=TRUE), type="l",col=5)
lines(1:12, colMeans(mape22,na.rm=TRUE), type="l",col=6)
lines(1:12, colMeans(mape23,na.rm=TRUE), type="l",col=7)
lines(1:12, colMeans(mape31,na.rm=TRUE), type="l",col=8)
lines(1:12, colMeans(mape32,na.rm=TRUE), type="l",col=9)
lines(1:12, colMeans(mape33,na.rm=TRUE), type="l",col=10)
legend("topleft", legend=c("bu-ARIMA","bu-ETS","bu-RW", "comb-ARIMA", "comb-ETS", "comb-RW", "tdgsa-ARIMA", "tdgsa-ETS", "tdgsa-RW"), col=2:10, lty=1, cex=0.5)
```

