---
title: "Product Demand Forecast-Monthly Model"
author: "Microsoft"
output: 
    rmarkdown::html_vignette:
        toc: true

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```

# Introduction
In this documentation, we are trying to build a monthly demand forecast model for total booking quantity data. We implement the advanced analysis process with R in the principle of step-by-step. 

## Total booking quantity data
Let's load the data and take a brief glimpse of it. 
```{r, message=FALSE}
##########################################################################
## Set directory
##########################################################################
setwd("C:/Users/zhouf/Documents/Rproject")
data.path <- "C:/Users/zhouf/Documents/Rproject"
```

```{r, message=FALSE, warning=FALSE, error=FALSE}
##########################################################################
## Read data
##########################################################################
library("readxl")
library("stringr")
rawData <- read_excel("Data/After scramble.xls")
# clean names up a bit
names(rawData) <- c("year", "month", "booking_qty", "sell_through_qty", "doi", "cogs", "casd")
# delete the first 6 rows
data <- rawData[-seq(1:6), ]
dim(data) # 108 7
str(data)
head(data)
```

## Model building

Let's visualize the monthly time series in the data.

```{r, message=FALSE, warning=FALSE, error=FALSE}
##########################################################################
## Time series analysis
##########################################################################
library(tseries)
library(forecast)
library(fpp)

# visualize monthly time series 
bts_mon <- ts(data[, 3:7], frequency=12, start=c(2006, 1))
head(bts_mon)
plot(bts_mon, xlab="Year", main="Monthly total booking quantity and features")

# partition data into train and test data set
train_mon <- window(bts_mon, start=c(2006, 1), end=c(2013, 12))
test_mon <- window(bts_mon, start=c(2014, 1), end=c(2014, 12))

# create training formula
labelVar = "booking_qty"
featureVars = colnames(train_mon)
featureVars = featureVars[which(featureVars!=labelVar)]
formula = as.formula(paste(paste(labelVar,"~"),paste(featureVars,collapse="+")))
```

### Regression model
```{r, message=FALSE, warning=FALSE, error=FALSE}
library(MASS)
fit <- lm(formula, data=train_mon)
fit
stepfit <- stepAIC(fit, direction="both")
stepfit$anova # display results 
fcast <- forecast(stepfit, newdata=as.data.frame(test_mon), h=12)
accuracy(fcast, test_mon[, 1])
fit <- lm(booking_qty ~ sell_through_qty + casd, data=train_mon)
fit
fcast <- forecast(fit, newdata=as.data.frame(test_mon), h=12)
fcast <- ts(fcast$mean, frequency=12, start=c(2014, 1))
accuracy(fcast, test_mon[, 1])
plot(fcast, type="l", col=2,
     xlab="Year", ylab="Total Booking Qty",
     main="Forecasts from regression",
     ylim=c(14000, 18000))
lines(test_mon[, 1], type="l", col=3)
legend("topleft",legend=c("Forecast","True"),col=2:3,lty=1)
```


### Support vector machine
```{r, message=FALSE, warning=FALSE, error=FALSE}
library(e1071)
fit <- svm(booking_qty ~ sell_through_qty + casd, data=train_mon)
fit
fcast <- predict(fit, newdata=test_mon)
fcast <- ts(fcast, frequency=12, start=c(2014, 1))
accuracy(fcast, test_mon[, 1])
plot(fcast, type="l", col=2,
     xlab="Year", ylab="Total Booking Qty",
     main="Forecasts from support vector machine",
     ylim=c(14000, 18000))
lines(test_mon[, 1], type="l", col=3)
legend("topleft",legend=c("Forecast","True"),col=2:3,lty=1)
```

### Regression forest (RevoScaleR)
```{r, message=FALSE, warning=FALSE, error=FALSE}
fit <- rxDForest(booking_qty ~ sell_through_qty + casd, data=train_mon, reportProgress=0)
fit
fcast <- rxPredict(fit, data=as.data.frame(test_mon), overwrite=TRUE)
fcast <- ts(fcast, frequency=12, start=c(2014, 1))
accuracy(fcast, test_mon[, 1])
plot(fcast, type="l", col=2,
     xlab="Year", ylab="Total Booking Qty",
     main="Forecasts from ScaleR regression forecast",
     ylim=c(14000, 18000))
lines(test_mon[, 1], type="l", col=3)
legend("topleft",legend=c("Forecast","True"),col=2:3,lty=1)
```

### Regression with ARIMA errors

```{r, message=FALSE, warning=FALSE, error=FALSE}
tsdisplay(diff(train_mon[, 1], 12))
tsdisplay(diff(diff(train_mon[, 1], 12)))
fit <- auto.arima(train_mon[,1], d=1, D=1, xreg=train_mon[, c(2,5)])
fit
fcast <- forecast(fit, xreg=test_mon[, c(2,5)], h=12)
accuracy(fcast, test_mon[, 1])
plot(fcast, xlab="Year", main="Forecasts from regression with ARIMA errors")
```

### Cross-validation
Next, we use cross validation to compare performance of different classes of models.

```{r, message=FALSE, warning=FALSE, error=FALSE}
# use time series cross validation to compare model performance
library(fpp)

k <- 84 # minimum data length for fitting a model
n <- dim(bts_mon)[1]
st <- tsp(bts_mon)[1]+(k-2)/12
mape1 <- mape2 <- mape3 <-mape4 <- matrix(NA,n-k,12)
for(i in 1:(n-k))
{
  xshort <- window(bts_mon, end=st + i/12)
  xnext <- window(bts_mon, start=st + (i+1)/12, end=st + (i+12)/12)
  fit1 <- lm(booking_qty ~ sell_through_qty + casd, data=xshort)
  fcast1 <- forecast(fit1, newdata=as.data.frame(xnext), h=12)
  fit2 <- svm(booking_qty ~ sell_through_qty + casd, data=xshort)
  fcast2 <- predict(fit2, newdata=xnext)
  fcast2 <- ts(fcast2, frequency=12, start=st + (i+1)/12, end=st + (i+12)/12)
  fit3 <- rxDForest(booking_qty ~ sell_through_qty + casd, data=xshort, reportProgress=0)
  fcast3 <- rxPredict(fit3, data=as.data.frame(xnext), overwrite=TRUE)
  fcast3 <- ts(fcast3, frequency=12, start=st + (i+1)/12, end=st + (i+12)/12)
  #fit4 <- auto.arima(xshort[, 1], d=1, D=1, xreg=xshort[, c(2,5)])
  #fcast4 <- forecast(fit4, xreg=xnext[, c(2,5)], h=12)
  mape1[i,1:dim(xnext)[1]] <- accuracy(fcast1, xnext[, 1])[2, "MAPE"]
  mape2[i,1:dim(xnext)[1]] <- accuracy(fcast2, xnext[, 1])[1, "MAPE"]
  mape3[i,1:dim(xnext)[1]] <- accuracy(fcast3, xnext[, 1])[1, "MAPE"]
  #mape4[i,1:dim(xnext)[1]] <- accuracy(fcast4, xnext[, 1])[2, "MAPE"]
}

plot(1:12, colMeans(mape1,na.rm=TRUE), type="l", col=2, 
     xlab="horizon", ylab="MAPE", ylim=c(0, 20))
lines(1:12, colMeans(mape2,na.rm=TRUE), type="l",col=3)
lines(1:12, colMeans(mape3,na.rm=TRUE), type="l",col=4)
#lines(1:12, colMeans(mape4,na.rm=TRUE), type="l",col=5)
#legend("topleft",legend=c("Linear","SVM","rxDForest","Reg with ARIMA"),col=2:5,lty=1,cex=0.5)
legend("topleft",legend=c("Linear","SVM","rxDForest"),col=2:4,lty=1,cex=0.5)
```




