---
title: "Product Demand Forecast-Monthly Model"
author: "Microsoft"
output: 
    rmarkdown::html_vignette:
        toc: true

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```

# Introduction
In this documentation, we are trying to build a monthly demand forecast model for total booking quantity data by using various regression models. When building the model, we include lagged total booking quantity as features into the model.

## Total booking quantity data
Let's load the data and take a brief glimpse of it. 
```{r, message=FALSE}
##########################################################################
## Set directory
##########################################################################
setwd("C:/Users/zhouf/Documents/Rproject")
data.path <- "C:/Users/zhouf/Documents/Rproject"
```

```{r, message=FALSE, warning=FALSE, error=FALSE}
##########################################################################
## Read data
##########################################################################
library("readxl")
library("stringr")
rawData <- read_excel("Data/After scramble.xls")
# clean names up a bit
names(rawData) <- c("year", "month", "booking_qty", "sell_through_qty", "doi", "cogs", "casd")
# delete the first 6 rows
data <- rawData[-seq(1:6), ]
dim(data) # 108 7
str(data)
head(data)
```

## Data cleaning and feature engineering
```{r, message=FALSE, warning=FALSE, error=FALSE}
###############################################################################################
## Data cleaning and feature engineering
###############################################################################################
library(tseries)
library(forecast)
library(fpp)

#------ identify lag --------------# 
y <- ts(data[, 3], frequency=12, start=c(2006, 1))
plot(y, xlab="Year", main="Monthly total booking quantity")
par(mfrow=c(1,2))
Acf(y, main="Total")
Pacf(y, main="Total") # lag 1, 7 , 12

#------ feature engineering ------------#
# lag features
shift = function(lag, x){c(rep(NA, lag), head(x,-lag))}
shift = Vectorize(shift, vectorize.args = "lag")

addlags = function(lags, df, var){
  res = shift(lags, x=df[,var])
  colnames(res) = paste("lag", lags, sep = "")
  return(cbind(df,res))
}

nlags = c(1, 7, 12)

newdata= addlags(nlags, data, var = "booking_qty")

head(newdata)

data <- newdata
```

## Data split
```{r, message=FALSE, warning=FALSE, error=FALSE}
###############################################################################################
## Split data into training and testing sets
###############################################################################################
## Data for training
out <- data
out.length <- dim(out)[1]
test.length <- 24
train.length <- out.length - test.length

train <- data.frame(out[1:train.length, ])
train <- train[complete.cases(train), ]
#train <- subset(train, select=-c(year,month))

## Data for testing
test  <- data.frame(out[(train.length+1):(train.length+test.length), ])
```

## Regression forest model
```{r, message=FALSE, warning=FALSE, error=FALSE}
###############################################################################################
## Train model
###############################################################################################
#create training formula
labelVar = "booking_qty"
featureVars = rxGetVarNames(train)
featureVars = featureVars[which((featureVars!=labelVar)&(featureVars!="year")&(featureVars!="month"))]
formula = as.formula(paste(paste(labelVar,"~"),paste(featureVars,collapse="+")))

#train regression forest model
regForest = rxDForest(formula, data = train, reportProgress=0)  
regForest
```

```{r, message=FALSE, warning=FALSE, error=FALSE}
###############################################################################################
## Score model
###############################################################################################
pred <- rxPredict(regForest, data=test, overwrite=TRUE)
result <- data.frame(year=test$year, 
                     month=test$month,
                     booking_qty=test$booking_qty, 
                     booking_qty_pred=pred)
head(result)
```

```{r, message=FALSE, warning=FALSE, error=FALSE}
###############################################################################################
## Evaluate model
###############################################################################################
MAPE <- mean(abs(result$booking_qty_Pred - result$booking_qty)/result$booking_qty)
MAPE
```

## Cross-validation to compare three regression models with same lagged and non-lagged features
Next, we use cross validation to compare performance of different classes of models.

### Models with all features
```{r, message=FALSE, warning=FALSE, error=FALSE}
###############################################################################################
## Time series cross validation to compare model performance
###############################################################################################
library(fpp)
library(e1071)

# model including all features
bts_mon <- ts(data[, 3:10], frequency=12, start=c(2006, 1))
head(bts_mon)
k <- 84 # minimum data length for fitting a model
n <- dim(bts_mon)[1]
st <- tsp(bts_mon)[1]+(k-2)/12
mape1 <- mape2 <- mape3 <- matrix(NA,n-k,12)
for(i in 1:(n-k))
{
  xshort <- window(bts_mon, end=st + i/12)
  xnext <- window(bts_mon, start=st + (i+1)/12, end=st + (i+12)/12)
  fit1 <- lm(formula, data=as.data.frame(xshort))
  fcast1 <- forecast(fit1, newdata=as.data.frame(xnext), h=12)
  fit2 <- svm(formula, data=xshort)
  fcast2 <- predict(fit2, newdata=xnext)
  fcast2 <- ts(fcast2, frequency=12, start=st + (i+1)/12, end=st + (i+12)/12)
  fit3 <- rxDForest(formula, data=xshort, reportProgress=0)
  fcast3 <- rxPredict(fit3, data=as.data.frame(xnext), overwrite=TRUE)
  fcast3 <- ts(fcast3, frequency=12, start=st + (i+1)/12, end=st + (i+12)/12)
  mape1[i,1:dim(xnext)[1]] <- accuracy(fcast1, xnext[, 1])[2, "MAPE"]
  mape2[i,1:dim(xnext)[1]] <- accuracy(fcast2, xnext[, 1])[1, "MAPE"]
  mape3[i,1:dim(xnext)[1]] <- accuracy(fcast3, xnext[, 1])[1, "MAPE"]
}

plot(1:12, colMeans(mape1,na.rm=TRUE), type="l", col=2, 
     xlab="horizon", ylab="MAPE", ylim=c(0, 20))
lines(1:12, colMeans(mape2,na.rm=TRUE), type="l",col=3)
lines(1:12, colMeans(mape3,na.rm=TRUE), type="l",col=4)
legend("topleft",legend=c("Linear with lags","SVM with lags","rxDForest with lags"),col=2:4,lty=1,cex=0.5)
```

### Models without quaterly features
```{r, message=FALSE, warning=FALSE, error=FALSE}
labelVar = "booking_qty"
featureVars = rxGetVarNames(data)
featureVars = featureVars[which((featureVars!=labelVar)&(featureVars!="year")&(featureVars!="month")&(featureVars!="doi")&(featureVars!="cogs"))]
formula = as.formula(paste(paste(labelVar,"~"),paste(featureVars,collapse="+")))

k <- 84 # minimum data length for fitting a model
n <- dim(bts_mon)[1]
st <- tsp(bts_mon)[1]+(k-2)/12
mape1 <- mape2 <- mape3 <- matrix(NA,n-k,12)
for(i in 1:(n-k))
{
  xshort <- window(bts_mon, end=st + i/12)
  xnext <- window(bts_mon, start=st + (i+1)/12, end=st + (i+12)/12)
  fit1 <- lm(formula, data=as.data.frame(xshort))
  fcast1 <- forecast(fit1, newdata=as.data.frame(xnext), h=12)
  fit2 <- svm(formula, data=xshort)
  fcast2 <- predict(fit2, newdata=xnext)
  fcast2 <- ts(fcast2, frequency=12, start=st + (i+1)/12, end=st + (i+12)/12)
  fit3 <- rxDForest(formula, data=xshort, reportProgress=0)
  fcast3 <- rxPredict(fit3, data=as.data.frame(xnext), overwrite=TRUE)
  fcast3 <- ts(fcast3, frequency=12, start=st + (i+1)/12, end=st + (i+12)/12)
  mape1[i,1:dim(xnext)[1]] <- accuracy(fcast1, xnext[, 1])[2, "MAPE"]
  mape2[i,1:dim(xnext)[1]] <- accuracy(fcast2, xnext[, 1])[1, "MAPE"]
  mape3[i,1:dim(xnext)[1]] <- accuracy(fcast3, xnext[, 1])[1, "MAPE"]
}

plot(1:12, colMeans(mape1,na.rm=TRUE), type="l", col=2, 
     xlab="horizon", ylab="MAPE", ylim=c(0, 20))
lines(1:12, colMeans(mape2,na.rm=TRUE), type="l",col=3)
lines(1:12, colMeans(mape3,na.rm=TRUE), type="l",col=4)
legend("topleft",legend=c("Linear with lags","SVM with lags","rxDForest with lags"),col=2:4,lty=1,cex=0.5)
```




